A few stickleback can transport adaptive alleles to new lakes

### 2019-04-24 Stickleback_Paper.tex
updated file and added Thom's MS

\title{A few stickleback suffice to transport adaptive alleles to new lakes}
\author{Jared Galloway, William A. Cresko, and Peter Ralph}
\maketitle

@@

Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being contected by recurrent gene flow. These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases smaller freshwater populations have been established ``de novo'' from marine colonists. Previous results show that a similar freshwater phenotype often evolves in independent populations soon after colonization using largely the same standing genetic variation, even in very distant geographic regions. Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation. In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in ocean populations. This process occurs over a realistic range of intermediate rates of gene flow that match previous empirical population genomic studies in stickleback. Examining these simulations more deeply reveals that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load but an increase probability of success for population genomic scans for locally adapted alleles. We also find that the original generation of marine individuals that colonized the lakes is most often the sole source of most freshwater adapted alleles. These simulations provide deeper context for existing studies of stickleback evolutionary genomics, and provide guidance for future empirical studies in this model. More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.

%\plr{say transporter hypothesis?}
%\plr{talk about linkage and haplotypes}

@@

The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations (\textbf{cite}). While this model has been very useful heuristically for guiding studies in the field and lab (\textbf{cite}), it is now rightfully understood as incomplete for many species in nature that have more complicated population structures (\textbf{cite}). In particular, empirical studies have increasingly pointed to the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats and experiencing diverse selective pressures while still exhibiting significant gene flow (\textbf{cite}). A growing number of empirical studies have identified instances of convergent evolution using SGV \citep{nelson2017ancient,Schrider2017,Barrett2008}. However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during invasion and adaptation to new environments (\textbf{cite}). It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV. While emprical studies in the field and lab are essential, the use of simulation models of specific systems can greatly advanced our understanding of the adaptation genetic dynamics using SGV.

@@

The threespine stickleback fish has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes (\textbf{cite}) and the much more extensive use of SGV than once thought (\textbf{cite}). The ancestral marine form of stickleback has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}). While geographic isolation prevents direct migration among freshwater populations, they often evolve similar (but not exactly the same) phenotypes (\textbf{cite}). The most recent evolutionary genomic studies on stickleback document that while adaptation to freshwater can occur over decades (\textbf{cite}), the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years (\textbf{cite}). Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations (\textbf{cite}). For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska (\textbf{cite}). Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years (\textbf{cite}). Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous studies of freshwater stickleback populations thousands of years old (\textbf{cite}).

But how can evolution occur at such a rapid pace in this and other species? Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake (\textbf{cite}). Even more improbable is having divergence cluster in such similar genomic regions across independent populations. An alternative hypothesis, supported by the most recent population genomic analyses, is that the marjority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations. In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter hypothesis'' to describe the process of alleles benefical in freshwater populations are maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation into new freshwater habitats. The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates (\textbf{cite}). While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere (\textbf{cite}). More recently, genome-wide haplotype analyses have provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}. These data support the transporter hypothesis, but a number of questions remain. What are the actual population sizes, migration rates, and fitness differentials consistent with this hypothesis? How many differentially selected alleles exist, how many are used at any one time, and how are they arranged within the genome?

Here we use forward simulations incorporating linkage to model the stickleback metapopulation and address these questions. We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds. Because we are documenting the entire genealogy of all alleles, we can determine the distribution and abundance of haplotypes across all in silico marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population. In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation. We find that only a few stickleback migrants per lake per generation are sufficient to maintain SGV in the ocean, allowing new freshwater populations can adapt in only tens of generations. We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers. Finally, we find that this rapid, local adaptation can only occur over an intermediate range of levels of gene flow. Low gene flow prevents variation from being transported, while high gene flow prevents local adaptation. Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.

@@

To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} 
to implement forwards-time simulation of populations of individuals with explicitly represented genomes
in which selection acted upon a single continuous quantitative trait.
The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography,
and included divergent selection in the two habitats which had substantial spatial structure.
Certain aspects of the model remain simplistic due to computational constraints.
Possibly the most important caveat is that simulated population sizes 
are much smaller than the census size of the threespine stickleback populations observed in nature
(see the Discussion for more on this).

\paragraph{Habitat and geography}
Simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures,
each with 5,000 diploid individuals.
The arrangement of these habitats, depicted in Figure~\ref{fig:Geo},
roughly models a set of freshwater habitats along a stretch of coastline. 
The marine habitat is a continuous, one-dimensional range of length 25 units,
while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''),
each connected to the marine habitat at regularly spaced intervals
(positions $i - 1/2$ for $1 \le i \le 25$).

Divergent selection is mediated by a single quantitative trait
with different optima in marine and freshwater habitats.
This situation roughly models the cumulative effect of the various phenotypes 
thought to be under divergent selection between the habitats,
such as armor morphology, body size, craniofacial variation and opercle shape.
The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively,
and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$
is determined by a Gaussian kernel with standard deviation 15, i.e.,
\begin{align*}
    f(x_\text{ind}; x_\text{opt})
    &=
    \exp\left\{
        \frac{1}{2}
            \left(
            \frac{x_\text{ind}-x_\text{opt}}{15}
            \right)^2
        \right\} .
\end{align*}
In short, the difference between an individual's trait value and the optimum determines that individual's fitness. 
% Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.
We chose the difference between optima and strength of stabilizing selection in each habitat
so that 
(a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,
and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.

@@

\paragraph{Genetic architecture of the trait}
Each individual carries two linear chromosomes, each of size $10^8$ loci.
Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation
in ten regions of $10^5$ loci each,
spread evenly along the chromosome.
Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). 
Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. 
Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. 
Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value 
$x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, 
where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$.
Subsequent mutations at the same locus replace the previous allele.
We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence}
to add neutral mutations across the entire genome at a rate of $10^{-7}$ per locus per generation, 
as described in \citet{kelleher2018efficient}.

\paragraph{Population dynamics}
We use SLiM to simulate a Wright--Fisher population with non-overlapping generations 
and a fixed population size of 5,000 diploid individuals in each habitat.
Each generation, the two parents of each new offspring are chosen proportional to their fitness 
(all individuals are hermaphroditic),
and the contributing genomes are produced by Poisson recombination 
with an average of one crossover per chromosome per generation 
($10^{-7}$ per locus per generation). 
Since the total population across \emph{all} 25 lakes is fixed at 5000,
and the Wright--Fisher model assumes unrealistic global population regulation,
we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation.
to do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, 
so that the mean fitnesses of all lakes are equal before selection happens.

As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat
as well as between the marine habitat and the lakes,
with a lake--ocean migration rate denoted $m$.
All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
Each new individual in each habitat 
has parents from the other habitat with probability $m$ 
(in which case we call it a ``migrant''),
and parents from the same habitat with probability $1-m$.
% how do we get nonmigrants?
The first parent of each nonmigrant individual in the freshwater habitat
is chosen from the freshwater habitat proportional to fitness, 
and a mate is chosen from the same lake as the first, also proportional to fitness. 
The offspring then lives in the same lake as the parents.
Parents for each nonmigrant marine individual are chosen similarly:
first, a single parent is chosen proportionally to fitness in the marine habitat,
and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. 
Concretely, if the first parent is marine individual $i$, 
then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$,
where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, 
and reflected to stay within the habitat range.
% what about migrants?
Parents for each freshwater migrant are chosen in the same way
as for nonmigrant marine individuals,
and are assigned to the lake nearest to the position of the first marine parent.
Similarly, parents for each marine migrant are both chosen from the same lake as before,
and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.

\paragraph{New lakes} 
To study how marine-derived populations adapt after colonizing newly appearing freshwater habitats,
we introduce a new set of 25 lakes after 100,000 generations.
These new lakes are populated with marine individuals to emulate a freshwater lake being colonized by
oceanic stickleback.
This creates two \emph{sets} of lakes, which along with the marine population have a total of 15,000 individuals.
Since this introduction of lakes doubles the number of lake-to-marine immigrants, 
the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.

@@

The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. 
To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome.
Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:
\begin{enumerate}
    \item \emph{a ``De novo" allele:}
        deriving from a new mutation that occurred in a new lake.
    \item \emph{a ``Migrant" allele:} 
        deriving from a migrant not in the initial generation that colonized the lake
    \item \emph{a ``Captured" allele:}
        present in initial colonists of the new lake, 
        and both common (above 50\%) in the original lakes,
        and uncommon (below 50\%) in the ocean.
    \item \emph{a ``Marine" allele:} 
        present in initial colonists of the new lake, 
        but not a ``captured'' allele.
\end{enumerate}
The proportion of trait-affecting alleles in new lakes that fall in these categories
measures the degree to which selection in the new environments made use of
(1) new mutation, 
(2) post-colonization migration, 
(3) standing variation at migration--selection balance, and
(4) standing variation at mutation--selection balance.

@@

As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations
with the exception of the highest gene flow at which half of each population was composed of migrants.
Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp}
show how mean trait values in freshwater and marine populations diverged over time,
until the trait means were close to the optimal values in each habitat. 
Establishment of new alleles in the lakes is visible 
in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value
which move the trait by an amount of order 1 every few hundred generations.
%Trait variation within each population was small compared to the difference between populations.
At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites 
(about 10 that shift the trait in each direction)
were responsible for most of the adaptive differences between freshwater and marine habitats.
As expected, increasing migration rate decreased differentiation between habitats:
as seen in Figure~\ref{fig:Fst}, 
$F_{ST}$ between marine and freshwater habitats 
at neutral sites steadily declines as migration increases.

@@

The numbers in Figure~\ref{fig:MPFAI} suggest that
the dramatic increase in speed of local adaptation we observed above occurs because
higher gene flow between populations allows sharing of freshwater alleles between populations.
%however, it is not enough for the marine population to hold bits and peices from independently formed haplotypes,
%For selection to act on SGV on the scale we observe in nature, the gene flow 
%initial generation to be able to rebuild 
We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele
in each individual in the new lakes, as defined in the Methods.
Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo''} mutation.
As gene flow increases, a larger fraction of adaptive alleles 
derive from pre-existing variation in the marine population at the time of introduction.
In other words, 
greater mixing at higher migration rates allows lakes to share alleles
instead of developing their own genetic basis of adaptation.

@@

At first glance, this suggests that genome scans for local adaptation based purely on measures of differentiation 
will only be successful given enough migration between habitats. 
But how many of these peaks are actually underlying trait differences that form the basis for local adaptation?
To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates 
that would be obtained by an $F_{ST}$ cutoff 
that declared everything above a certain value to be a causal locus.
\plr{need to say what true and false positive windows are}
What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake, per generation
are the only rates of dispersal which provide reliable true negatives when observing peaks.
This means that a large percentage of peaks past an $F_{ST}$ cutoff of $0.05$ 
actually lie on top of regions which could impact trait differences. 
Unfortunately, the low statistical power for all rates of $M$ would seem to suggest
that many regions which impact trait differences do not appear as $F_{st}$ peaks. 
This may imply that genome scans for local adaptation based purely on measures of differentiation may not be as reliable as previously thought.
To summarize, our study find that with high rates of dispersal between population, $F_{st}$ peaks may be a good indication of causal loci.
However, this is a one way implication, meaning even with high dispersal, not all causal loci will appear as $F_{st}$ peaks.

@@

Suppose a new allele enters a lake, either by migration or mutation. 
If, when it is rare but present in $n$ copies, it has fitness advantage $s$ 
-- i.e., the expected number of copies in the next generation is $(1+s)n$ 
-- then the probability that it escapes demographic stochasticity to become common in the population 
is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical}
(assuming Poisson reproduction, as we roughly have here).
If the current population all differ from the optimum trait by $z$, 
and the allele has effect size $-u$ in heterozygotes, 
then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$. 
This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.

>>>>

\title{A few stickleback can transport adaptive alleles to new lakes}
\author{Jared Galloway, William A. Cresko, and Peter Ralph}
\maketitle

@@

Threespine stickleback provide a striking example of local adaptation to divergent habitats while still being contected by recurrent gene flow. These small fish occur in marine and freshwater habitats throughout the Northern Hemisphere, and in numerous cases smaller freshwater populations have been established ``de novo'' from marine colonists. Previous results show that a similar freshwater phenotype often evolves in independent populations soon after colonization using largely the same standing genetic variation, even in very distant geographic regions. But how many stickleback migrants are needed to capture a sufficient amount of standing variation to fuel adaptation? Here we use genealogical simulations to determine the levels of gene flow that best match observed patterns of allele sharing among habitats in stickleback, and to better understand how gene flow and local adaptation in large metapopulations determine the speed of adaptation and re-use of standing genetic variation. In our simulations we find that repeated adaptation uses a shared set of alleles that are maintained at low frequency by migration-selection balance in ocean populations. This process occurs over a realistic range of intermediate levels of gene flow that match previous empirical population genomic studies in stickleback. Examining these simulations more deeply reveals that low gene flow leads to slow, independent adaption to district habitats, whereas high gene flow leads to significant mutation load but an increased probability of successful population genomic scans for locally adapted alleles. We also find that the original generation of marine individuals that colonized the lakes is most often the sole source of most freshwater adapted alleles. These simulations provide deeper context for existing studies of stickleback evolutionary genomics, and provide guidance for future empirical studies in this model. More broadly, our results support existing theory of local adaptation but extend it by more completely documenting the genealogical history of adaptive alleles in a metapopulation.

@@

The canonical model for the genetics of adaptation has long been the sequential fixation of new mutations (\textbf{cite}). While it has been very useful heuristically for guiding studies in the field and lab (\textbf{cite}), this model is now rightfully understood as incomplete for many species in nature that have more complicated population structures (\textbf{cite}). In particular, empirical studies have increasingly pointed to the need to more deeply incorporate standing genetic variation (SGV) into adaptation dynamics for metapopulations inhabiting an array of habitats and experiencing diverse selective pressures while still exhibiting significant gene flow (\textbf{cite}). A growing number of empirical studies have identified instances of convergent evolution using SGV \citep{nelson2017ancient,Schrider2017,Barrett2008}. However, it is still not clear how variation in evolutionary processes - such as gene flow, recombination, selection and mutation -  can promote the maintenance and re-use of SGV, particularly during invasion and adaptation to new environments (\textbf{cite}). It is similarly unclear whether variation in these evolutionary processes can determine the genetic architecture of evolving traits via SGV. While emprical studies in the field and lab are essential, the use of simulation models of specific systems can greatly advanced our understanding of the adaptation genetic dynamics using SGV.

@@

The threespine stickleback fish has provided some of the earliest data in a recent flood showing the heterogeneous nature of divergence across genomes (\textbf{cite}) and the much more extensive use of SGV than once thought (\textbf{cite}). The ancestral marine form of stickleback has given rise to millions of independently derived freshwater populations in recently de-glaciated regions around the Northern Hemisphere (\textbf{cite}). While geographic isolation often prevents direct migration among freshwater populations, they frequently evolve similar (but not exactly the same) phenotypes (\textbf{cite}). The most recent evolutionary genomic studies on stickleback document that while adaptation to freshwater can occur over decades (\textbf{cite}), the overall dynamic of local adaptation to marine and freshwater habitats has been occurring for millions of years (\textbf{cite}). Impressively, independent local adaptation of marine individuals to freshwater environments has been observed to take place in just tens of generations (\textbf{cite}). For example, in 1964 the Great Alaskan Earthquake caused an uplift of many islands and coastal regions throughout the Gulf of Alaska (\textbf{cite}). Studies of stickleback populations on uplifted Middleton Island showed that newly created freshwater ponds were invaded by the surrounding marine population of stickleback which evolved the freshwater syndrome of phenotypes in less than 50 years (\textbf{cite}). Amazingly, nearly a quarter of the genome diverged significantly in this same time but - with very little change in the rest of the genome - in a pattern that mirrored previous studies of freshwater stickleback populations thousands of years old (\textbf{cite}).

But how can evolution occur at such a rapid pace in this and other species? Waiting for new mutations to arise in each lake or pond would take much longer than the decades since the Alaskan earthquake (\textbf{cite}). Even more improbable is having divergence cluster in such similar genomic regions across independent populations. An alternative hypothesis, supported by the most recent population genomic analyses, is that the marjority of alleles important for freshwater adaptation are maintained in the marine individuals due to recurrent gene flow from freshwater back in to marine populations. In 2009 \citet{schluter2009genetics} proposed a conceptual model they termed the ``transporter hypothesis'' to describe the process of alleles benefical in freshwater populations are maintained at migration-selection balance in the larger oceanic population and thus available for subsequent adaptation into new freshwater habitats. The first clear example of the global reuse of SGV in stickleback was the gene \textit{eda} which has been shown to be an important regulator for the number of bony lateral plates (\textbf{cite}). While the low lateral plate version of this gene arose millions of years ago, it is found in much younger freshwater ponds around the Northern Hemisphere (\textbf{cite}). More recently, genome-wide haplotype analyses have provided evidence that \textit{most} regions of the genome that distinguish marine-freshwater genetic differences share this pattern \citep{nelson2017ancient}.

While this growing body of population genomic data on stickleback evolution support the transporter hypothesis, a number of questions remain. What are the actual population sizes, migration rates, and fitness differentials consistent with this hypothesis? How many differentially selected alleles exist, how many are used at any one time, and how are they arranged within the genome? A curious natural history observation undergirding many of these questions is the fact that some newly formed freshwater habitats, such as the ponds on Middleton Island, are quite small and presumably the number of initial marine migrants is few. Here we use forward simulations implementing in SLiM and incorporating linkage to model the stickleback metapopulation and address these questions. We ask how variation in amount of gene flow affects the genetic architecture of local adaptation to newly created freshwater ponds. Because we are documenting the entire genealogy of all alleles, we can determine the distribution and abundance of haplotypes across all in silico marine and freshwater populations to know the timing and proportion of potentially adaptive alleles that are actually utilized in each population. In addition, we can document how these adaptive alleles are distributed across the genomes, and as a consequence determine how this may affect the efficacy of genome scans for between-habitat differentiation. We find that only a few stickleback migrants per lake per generation are sufficient to maintain SGV in the ocean, allowing new freshwater populations can adapt in only tens of generations. We also find that continued gene flow of marine individuals into recently established lakes is not necessary for adaptation as the majority of alleles are present in the initial marine colonizers. Finally, we find that this rapid, local adaptation can only occur over an intermediate range of levels of gene flow. Low gene flow prevents variation from being transported, while high gene flow prevents local adaptation. Together these simulation findings enhance and extend the existing rich empirical work on stickleback evolutionary genomics.

@@

To explore these questions, we used SLiM \citep{haller2017slim,haller2018slim3} to implement forward-time simulation of populations of individuals with explicitly represented genomes in which selection acted upon a single continuous quantitative trait. The details of the model were motivated by current understanding of threespine stickleback evolutionary history and demography, and included divergent selection in the two habitats which had substantial spatial structure. Certain aspects of the model remain simplistic due to computational constraints. Possibly the most important caveat is that simulated population sizes are much smaller than the census size of the threespine stickleback populations observed in nature (see the Discussion for more on this).

\paragraph{Habitat and geography}

Simulations have two habitat types -- Marine and Freshwater -- defined by their selective pressures, each with 5,000 diploid individuals. The arrangement of these habitats, depicted in Figure~\ref{fig:Geo} roughly models a set of freshwater habitats along a stretch of coastline. The marine habitat is a continuous, one-dimensional range of length 25 units, while the freshwater habitat is divided into 25 discrete subpopulations (which we call ``lakes''), each connected to the marine habitat at regularly spaced intervals (positions $i - 1/2$ for $1 \le i \le 25$).

Divergent selection is mediated by a single quantitative trait with different optima in marine and freshwater habitats. This situation roughly models the cumulative effect of the various phenotypes thought to be under divergent selection between the habitats, such as armor morphology, body size, craniofacial variation and opercle shape. The optimal trait values in the marine and freshwater habitats are $+10$ and $-10$ respectively, and fitness of a fish with trait value $x_\text{ind}$ in a habitat with optimal value $x_\text{opt}$ is determined by a Gaussian kernel with standard deviation 15, i.e.,

\begin{align*}
    f(x_\text{ind}; x_\text{opt})
    &=
    \exp\left\{
        \frac{1}{2}
            \left(
            \frac{x_\text{ind}-x_\text{opt}}{15}
            \right)^2
        \right\} .
\end{align*}

In short, the difference between an individual's trait value and the optimum determines that individual's fitness. We chose the difference between optima and strength of stabilizing selection in each habitat so that (a) around 10 (diploid, homozygous) mutations were sufficient to move from one optimum to the other,and (b) well-adapted fish from one habitat would have low, but nonzero, fitness in the other habitat.
% Note that the scale we measure trait on is arbitrary as only the scale of trait values relative to mutation effect size is relevant.

@@

\paragraph{Genetic architecture of the trait}

Each individual carries two linear chromosomes, each of size $10^8$ loci. Mutations that can affect the trait under selection can occur at rate $10^{-10}$ per locus per generation in ten regions of $10^5$ loci each, spread evenly along the chromosome. Each mutation in these regions is either additive, completely recessive, or completely dominant (with equal probability). Effect sizes for these mutations are chosen randomly from an exponential distribution with mean $1/2$, either positive or negative with equal probability. Individual trait values ($x_\text{ind}$) are determined additively from the diploid genotypes. Concretely, an individual that is heterozygous and homozygous for mutations at sets of loci $H$ and $D$ respectively has trait value $x_\text{ind} = \sum_{i \in H} h_i s_i + \sum_{j \in D} s_j$, where $h_i$ and $s_i$ are the dominance coefficient and the effect size of the mutation at locus $i$. Subsequent mutations at the same locus replace the previous allele. We also used SLiM's ability to record \emph{tree sequences} \citep{haller2018treesequence} to add neutral mutations across the entire genome at a rate of $10^{-7}$ per locus per generation,  as described in \citet{kelleher2018efficient}.

\paragraph{Population dynamics}

We use SLiM to simulate a Wright--Fisher population with non-overlapping generations and a fixed population size of 5,000 diploid individuals in each habitat. Each generation, the two parents of each new offspring are chosen proportional to their fitness (all individuals are hermaphroditic), and the contributing genomes are produced by Poisson recombinationwith an average of one crossover per chromosome per generation ($10^{-7}$ per locus per generation). Since the total population across \emph{all} 25 lakes is fixed at 5000, and the Wright--Fisher model assumes unrealistic global population regulation, we normalize the fitnesses of each individual such that approximately 200 offspring are generated in each lake, each generation. To do this, we divide fitness values of each freshwater individual by the mean fitness in their lake, so that the mean fitnesses of all lakes are equal before selection happens.

As depicted in Figure~\ref{fig:Geo}, dispersal occurs both locally along the coastline in the marine habitat as well as between the marine habitat and the lakes,
with a lake--ocean migration rate denoted $m$.All individual dispersal events can be thought of as occurring at the juvenile stage in the life cycle of the simulation. 
Each new individual in each habitat has parents from the other habitat with probability $m$ (in which case we call it a ``migrant''),and parents from the same habitat with probability $1-m$.
% how do we get nonmigrants?
The first parent of each nonmigrant individual in the freshwater habitat is chosen from the freshwater habitat proportional to fitness, and a mate is chosen from the same lake as the first, also proportional to fitness. The offspring then lives in the same lake as the parents. Parents for each nonmigrant marine individual are chosen similarly:
first, a single parent is chosen proportionally to fitness in the marine habitat, and then a mate is chosen, also proportionally to fitness but re-weighted by a Gaussian function of the distance separating the two, with standard deviation $1/2$. Concretely, if the first parent is marine individual $i$, then marine individual $j$ is chosen as the mate with probability proportional to $f(x_j) \exp(-2d_{ij}^2)$, where $f(x_j)$ is the fitness of individual $j$ and $d_{ij}$ is the distance between the two locations. 
Finally, each new marine offspring is given a position displaced from the first parent's position by a random Gaussian distance with mean 0 and standard deviation 0.5, and reflected to stay within the habitat range.
% what about migrants?
Parents for each freshwater migrant are chosen in the same way as for nonmigrant marine individuals, and are assigned to the lake nearest to the position of the first marine parent. Similarly, parents for each marine migrant are both chosen from the same lake as before, and the offspring is given a spatial location in the marine habitat at the location of the parent's lake.

\paragraph{New lakes}

To study how marine-derived populations adapt after colonizing newly appearing freshwater habitats, we introduce a new set of 25 lakes after 100,000 generations. These new lakes are populated with marine individuals to emulate a freshwater lake being colonized by oceanic stickleback. This creates two \emph{sets} of lakes, which along with the marine population have a total of 15,000 individuals. Since this introduction of lakes doubles the number of lake-to-marine immigrants,the probability that a new marine individual has freshwater parents is $2m$ instead of $m$.

@@

The output tree sequence from each simulation allows us to explore the origin of the genetic basis of adaptation in the new lakes. To do this, we constructed the genealogical tree relating all extant chromosomes at each locus along the genome. Using these trees we classified each adaptive allele, in each genome in the new lakes at the time of adaptation, into four categories:

\begin{enumerate}
    \item \emph{a ``De novo" allele:}
        deriving from a new mutation that occurred in a new lake.
    \item \emph{a ``Migrant" allele:} 
        deriving from a migrant not in the initial generation that colonized the lake
    \item \emph{a ``Captured" allele:}
        present in initial colonists of the new lake, 
        and both common (above 50\%) in the original lakes,
        and uncommon (below 50\%) in the ocean.
    \item \emph{a ``Marine" allele:} 
        present in initial colonists of the new lake, 
        but not a ``captured'' allele.
\end{enumerate}

The proportion of trait-affecting alleles in new lakes that fall in these categories measures the degree to which selection in the new environments made use of
(1) new mutation, 
(2) post-colonization migration, 
(3) standing variation at migration--selection balance, and
(4) standing variation at mutation--selection balance.

@@

As shown in Figure~\ref{fig:MeanPhenotype}, local adaptation occurred in all simulations with the exception of the highest gene flow at which half of each population was composed of migrants. Figures \ref{fig:phenotype_time} and \ref{fig:phenotype_time_supp} show how mean trait values in freshwater and marine populations diverged over time, until the trait means were close to the optimal values in each habitat. Establishment of new alleles in the lakes is visible in the first few thousand generations of Figure \ref{fig:phenotype_time} as jumps in the mean trait value which move the trait by an amount of order 1 every few hundred generations.
%Trait variation within each population was small compared to the difference between populations.
At the lowest rate of gene flow, $M = 0.01$, differences at around 20 commonly polymorphic sites(about 10 that shift the trait in each direction) were responsible for most of the adaptive differences between freshwater and marine habitats. As expected, increasing migration rate decreased differentiation between habitats: as seen in Figure~\ref{fig:Fst}, $F_{ST}$ between marine and freshwater habitats at neutral sites steadily declines as migration increases.

@@

The numbers in Figure~\ref{fig:MPFAI} suggest that the dramatic increase in speed of local adaptation we observed above occurs because higher gene flow between populations allows sharing of freshwater alleles between populations.

%however, it is not enough for the marine population to hold bits and peices from independently formed haplotypes,
%For selection to act on SGV on the scale we observe in nature, the gene flow 
%initial generation to be able to rebuild

We confirmed this by using recorded tree sequences to identify the origin of each trait-affecting allele in each individual in the new lakes, as defined in the Methods. Figure~\ref{fig:Origin} shows that at the lowest rate of gene flow the majority of adaptive alleles are derived from \emph{``de novo''} mutation. As gene flow increases, a larger fraction of adaptive alleles derive from pre-existing variation in the marine population at the time of introduction. In other words, greater mixing at higher migration rates allows lakes to share alleles instead of developing their own genetic basis of adaptation.

@@

At first glance, this suggests that genome scans for local adaptation based purely on measures of differentiation will only be successful given enough migration between habitats. But how many of these peaks are actually underlying trait differences that form the basis for local adaptation? To quantify this, Figure \ref{fig:Power_FP} shows the power and false positive rates 
that would be obtained by an $F_{ST}$ cutoff that declared everything above a certain value to be a causal locus.

\plr{need to say what true and false positive windows are}

What we observe most plainly in this graph is that $M = 1.0$ and $M = 10.0$ migrants per lake, per generation are the only rates of dispersal which provide reliable true negatives when observing peaks. This means that a large percentage of peaks past an $F_{ST}$ cutoff of $0.05$ actually lie on top of regions which could impact trait differences. Unfortunately, the low statistical power for all rates of $M$ would seem to suggest that many regions which impact trait differences do not appear as $F_{st}$ peaks. This may imply that genome scans for local adaptation based purely on measures of differentiation may not be as reliable as previously thought. To summarize, our study find that with high rates of dispersal between population, $F_{st}$ peaks may be a good indication of causal loci. However, this is a one way implication, meaning even with high dispersal, not all causal loci will appear as $F_{st}$ peaks.

@@

Suppose a new allele enters a lake, either by migration or mutation. If, when it is rare but present in $n$ copies, it has fitness advantage $s$ -- i.e., the expected number of copies in the next generation is $(1+s)n$ -- then the probability that it escapes demographic stochasticity to become common in the population is approximately $2s$ \citep{lambert2006probability,haldane1927mathematical(assuming Poisson reproduction, as we roughly have here). If the current population all differ from the optimum trait by $z$, and the allele has effect size $-u$ in heterozygotes,  then the fitness advantage of the allele would be $s(u) = \exp(-\beta((z - u)^2) / \exp( - \beta z^2)) \approx 2 \beta z u$, where in our parameterization, $\beta = 1 / 450$. This tells us two things: (1) the rate of adaptation decreases as the population approaches the optimum, and (2) larger mutations (in the right direction) are more likely to fix.

\paragraph{New mutations}
